HTW Model e2

### 2024-02-24 Analysis/e1_test.qmd
rerender after reorganization; Results toc adjustments; remove multi-column and echo from indiv. results files.

```
## Analyses Strategy

@@

{{< include ../Misc/bmm_table.qmd >}}

## Results

### Testing Phase - No feedback.

@@

#### Deviation From Target Band

@@

```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 1
#| echo: fenced

@@

#### Discrimination between bands

@@

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1
#| echo: fenced

@@

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| layout-ncol: 1
#| echo: fenced
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands")

@@

#### E1 Results Discussion

@@

## References

::: {#refs}
:::

## Appendix

>>>>

```
### Analyses Strategy

@@

::: column-page-inset-right
{{< include ../Misc/bmm_table.qmd >}}
:::

### Results

#### Testing Phase - No feedback.

@@

##### Deviation From Target Band

@@

```{r}
#| label: tbl-e1-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 1

@@

##### Discrimination between bands

@@

```{r}
#| label: tbl-e1-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1

@@

```{r}
#| label: tbl-e1-bmm-vx
#| tbl-cap: "Experiment 1. Bayesian Mixed Model Predicting Vx as a function of condition (Constant vs. Varied) and Velocity Band"
#| tbl-subcap: ["Model fit to all 6 bands", "Model fit to 3 extrapolation bands"]
#| layout-ncol: 1
e1_vxBMM <- brm(vx ~ condit * bandInt + (1 + bandInt|id),
                        data=test,file=paste0(here::here("data/model_cache", "e1_testVxBand_RF_5k")),
                        iter=5000,chains=4,silent=0,
                        control=list(adapt_delta=0.94, max_treedepth=13))
GetModelStats(e1_vxBMM ) |> kable(escape=F,booktabs=T, caption="Fit to all 6 bands")

@@

### E1 Results Discussion

@@

### 2024-02-24 Analysis/e2_test.qmd
rerender after reorganization; Results toc adjustments; remove multi-column and echo from indiv. results files.

## Results

### Testing Phase - No feedback.

@@

```{r}
#| label: tbl-e2-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| echo: fenced

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied 
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

#### Discrimination between Velocity Bands

@@

```{r}
#| label: tbl-e2-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 2
#| echo: fenced

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant 
result$varied

@@

When examining discrimination ability using the model predicting raw x-velocity, the results were less clear than those of the absolute deviation analysis. The slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]) indicates that participants showed good discrimination between bands overall. However, the interaction term suggested this effect was not modulated by training condition (β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\]) Thus, while varied training may provide some advantage for accuracy, it does not appear to impair the ability to discriminate between velocity bands relative to constant training.

@@

@fig-e2-bmm-bx2 shows the distributions of estimated slopes relating velocity band to x velocity for each participant, ordered from lowest to highest within condition. Slope values are lower overall for varied training compared to constant training. Figure Xb plots the density of these slopes for each condition. The distribution for varied training has more mass at lower values than the constant training distribution. Both figures illustrate the model's estimate that varied training resulted in less discrimination between velocity bands, evidenced by lower slopes on average.

>>>>

### Results

#### Testing Phase - No feedback.

@@

```{r}
#| label: tbl-e2-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]

result <- test_summary_table(testE2, "dist","Deviation", mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl()
result$varied |> kbl()
# make kable table with smaller font size
# result$constant |> kbl(caption="Constant Testing - Deviation",booktabs=T,escape=F) |> kable_styling(font_size = 7)
```

@@

##### Discrimination between Velocity Bands

@@

```{r}
#| label: tbl-e2-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1

result <- test_summary_table(testE2, "vx","X Velocity" ,mfun = list(mean = mean, median = median, sd = sd))
result$constant |> kbl()
result$varied |> kbl()

@@

When examining discrimination ability using the model predicting raw x-velocity, the results were less clear than those of the absolute deviation analysis. The slope on Velocity Band (β = `r sc1$estimate`, 95% CrI \[`r sc1$conf.low`, `r sc1$conf.high`\]) indicates that participants showed good discrimination between bands overall. However, the interaction term suggested this effect was not modulated by training condition (β = `r intCoef1$estimate`, 95% CrI \[`r intCoef1$conf.low`, `r intCoef1$conf.high`\]) Thus, while varied training may provide some advantage for accuracy, both training conditions seem to have similar abilities to discriminate between velocity bands.

@@

@fig-e2-bmm-bx2 visually represents the distributions of estimated slopes relating velocity band to x velocity for each participant, ordered from lowest to highest within condition. Slope values are lower overall for varied training compared to constant training. Figure Xb plots the density of these slopes for each condition. The distribution for varied training has more mass at lower values than the constant training distribution. Both figures illustrate the model's estimate that varied training resulted in less discrimination between velocity bands, evidenced by lower slopes on average.

### 2024-02-24 Analysis/e3_test.qmd
add manuscripts
rerender after reorganization; Results toc adjustments; remove multi-column and echo from indiv. results files.

```{r}
#| label: tbl-e3-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]
#| layout-ncol: 2
#| echo: fenced

@@

```{r}
#| label: tbl-e3-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 2
#| echo: fenced

resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

>>>>

```{r}
#| label: tbl-e3-test-nf-deviation
#| tbl-cap: "Testing Deviation - Empirical Summary"
#| tbl-subcap: ["Constant Testing - Deviation", "Varied Testing - Deviation"]

@@

```{r}
#| label: tbl-e3-test-nf-vx
#| tbl-cap: "Testing vx - Empirical Summary"
#| tbl-subcap: ["Constant Testing - vx", "Varied Testing - vx"]
#| layout-ncol: 1
resultOrig <- test_summary_table(testE3 |> filter(bandOrder=="Original"), "vx","X Velocity", mfun = list(mean = mean, median = median, sd = sd))
resultOrig$constant 
resultOrig$varied

### 2024-02-24 Model/htw_model_e2.qmd
rerender after reorganization; Results toc adjustments; remove multi-column and echo from indiv. results files.

To more accurately assess the relative abilities of ALM and EXAM to capture important empirical patterns - we will now examine the predictions of both models for the subset of individual participants shown in @fig-htw-indv-pred. Panel A presents three varied and constant participants who demonstrated a reasonable degree of discrimination between the 6 velocity bands during testing.

-   \*\* comment on the different ways ALM can completely fail to mimic discrimination patterns (sbj. 35; sbj. 137),and on how it can sometimes partially succeed (sbj. 11; 14,74)

-   \*\* comment on how EXAM can somtimes mimic non-monotonic spacing between bands due to associative stregth from training (i.e. subject 47)

-   \*\* compare c values to slope parameters from the statistical models earlier in paper

```{r  fig.width=11, fig.height=10}
#| label: fig-htw-indv-pred
#| fig-cap: Model predictions alongside observed data for a subset of individual participants. A) 3 constant and 3 varied participants fit to both the test and training data. B) 3 constant and 3 varied subjects fit to only the trainign data. 
#| fig-height: 13
#| fig-width: 12

@@

#### Subjects with no clear best model
```{r}
#| fig-height: 6
#| fig-width: 12
vAlm <- c(129,192,105); cAlm <- c(101, 109,134)

post_dat_l |> filter(id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |> 
   mutate(x=as.factor(x), Resp=as.factor(Resp)) |>
  group_by(id,condit,Fit_Method,Model,Resp) |>
   mutate(flab=paste0("Subject: ",id)) |>
  ggplot(aes(x = Resp, y = val, fill=x)) + 
  stat_bar_sd + ggh4x::facet_nested_wrap(condit~flab, axes = "all",ncol=3) +
  labs(title="Subjects with no clear best model",
       y="X Velocity",fill="Target Velocity") +
   guides(fill = guide_legend(nrow = 1)) + 
  theme(legend.position = "bottom",axis.title.x = element_blank())

```{r}
#| fig-height: 8
#| fig-width: 12
#| 
vAlm <- c(29,70,68); cAlm <- c(73,49,128)

difAlm <-  post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=log(c), x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.5)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="c parameter - sbjs. with biggest diff. favoring ALM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(-11,-4))

vAlm <- c(23,155,184); cAlm <- c(119,85,175)

difExam <- post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=log(c), x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.1)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="c parameter - sbjs. with biggest diff. favoring EXAM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(-11,-4))

difAlm/difExam

vAlm <- c(29,70,68); cAlm <- c(73,49,128)

difAlm <-  post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=lr, x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.5)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="lr parameter - sbjs. with biggest diff. favoring ALM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4))+ylim(c(0,12))

vAlm <- c(23,155,184); cAlm <- c(119,85,175)

difExam <- post_dat_avg  |> filter(x==100,id %in% c(vAlm,cAlm), Fit_Method=="Test_Train") |>
    group_by(id, condit, Model, rank) %>%
      mutate(flab=paste0("Subject: ",id)) |>
    ggplot(aes(y=lr, x = id,col=Model)) + 
  stat_pointinterval(position=position_dodge(.1)) +
   # ggh4x::facet_nested_wrap(condit~flab, axes = "all",scales="free_y",ncol=3) + 
      labs(title="lr parameter - sbjs. with biggest diff. favoring EXAM") +
  theme(legend.title = element_blank(), legend.position="right",plot.title=element_text(hjust=.4)) +ylim(c(0,12))

difAlm/difExam
```

```{r fig.width=11, fig.height=13}

# 
# 
# full_comparison |> filter(Fit_Method=="Test_Train") |>
#   ungroup() |>
#   mutate(id = reorder(id, Error_difference)) %>%
#   ggplot(aes(y=id,x=Error_difference,fill=Best_Model))+
#   geom_col()+
#   ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")
#

# d <- testAvg |> left_join(full_comparison, by=c("id","condit")) |> filter(Fit_Method=="Test_Train")
# 
# 
# 
# d |> ggplot(aes(x=vb,y=vx,fill=condit)) + stat_bar + facet_wrap(Fit_Method~Best_Model2,ncol=2)
# 
# d |> 
#   group_by(condit,Fit_Method,Best_Model) |> 
#   mutate(nGrp2=n()) |>
#   filter(abs(Error_difference)>15) |> 
#   ggplot(aes(x=vb,y=dist,fill=condit)) + 
#   stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)
# 
# d |> group_by(condit,Fit_Method,Best_Model) %>% tally() |> mutate(n=n/6)
# 
# d |> group_by(condit,Fit_Method,Best_Model) |> filter(abs(Error_difference)>15) |> tally() |> mutate(n=n/6)
# 
# d |> group_by(condit,Fit_Method) |> mutate(m=mean(Error_difference), 
#                                               sd=sd(Error_difference), 
#                                               n=n()/6,se=sd/sqrt(n)) |>
#    group_by(condit,Fit_Method,Best_Model) |> 
#  # filter(abs(Error_difference)>(2.5*se)) |> 
#   ggplot(aes(x=vb,y=dist,fill=condit)) + 
#   stat_bar + facet_wrap(Fit_Method~Best_Model,ncol=2)
#

```{r fig.width=11, fig.height=13}
#| eval: false
#| include: false

posterior <- post_dat |> group_by(id, condit, Model, Fit_Method,rank) |> 
    slice_head(n=1) |> select(id,condit,Model,Fit_Method,rank,c,lr, mean_error)
head(posterior)

# Combine the MAP density values with your original posterior data frame
posterior_with_density <- left_join(posterior, posterior_densities, by = c("id", "condit", "Model", "Fit_Method"))

head(posterior_with_density)

posterior %>%
  group_by(id, condit, Model, Fit_Method) %>%
  do({
    data = .
    c_density = density(data$c)
    lr_density = density(data$lr)

    # Find the index of the MAP estimate in your density objects
    map_c_index = which.min(abs(c_density$x - data$c[data$rank == 1]))
    map_lr_index = which.min(abs(lr_density$x - data$lr[data$rank == 1]))

    data.frame(
      c_density_at_MAP = c_density$y[map_c_index],
      lr_density_at_MAP = lr_density$y[map_lr_index]
    )
  }) -> densities_at_map

# Combine the MAP density values with your original posterior data frame
posterior_with_density <- left_join(posterior, posterior_densities, by = c("id", "condit", "Model", "Fit_Method"))

head(posterior_with_density)

map_id <- post_dat |> group_by(id, condit, Model, Fit_Method) |> 
  filter(rank==1) |> slice_head(n=1) |> select(id,condit,Fit_Method,Model,c,lr,mean_error,rank) |>
  left_join(id_n,by=(c("id","condit")))

map_id <- map_id |> 
  mutate(n = fifelse(Fit_Method == "Test", nTest, fifelse(Fit_Method == "Train", nTrain, nTotal)),
         BIC = -2 * log(1/mean_error) + log(n)*2) |>
  group_by(id,Fit_Method) |>
  mutate(Best_Model=Model[which.min(BIC)])

bic_id <- map_id %>%
  group_by(id, condit, Fit_Method) %>%
  arrange(Model) %>%
  summarize(BIC_diff = diff(BIC),
            .groups = 'drop')

 bic_id |> filter(Fit_Method=="Test_Train") |>
     ungroup() |>
   mutate(id = reorder(id, BIC_diff),decreasing=FALSE) %>%
   ggplot(aes(y=id,x=BIC_diff))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")

bic_id |> group_by(condit,Fit_Method) |> summarise(mean(BIC_diff))

### To add to appendix

```{r}

post_tabs$agg_x_full |> flextable::tabulator(rows=c("Fit_Method","x"), columns=c("condit","Model"),
                       `X` = as_paragraph(mean_error)) |> as_flextable()

# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method,x) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   round_tibble(1) |> pander::pandoc.table()

# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |>
#   mutate(e2=abs(y-pred)) |>
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |>
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y1), pred=mean(pred1),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred)) |>
#   group_by(condit,Model,Fit_Method,x) |> 
#   summarise(y2=mean(y1), pred2=mean(pred1)) |> 
#   group_by(condit,Model,Fit_Method) |> 
#   summarise(y=mean(y2), pred=mean(pred2),mean_error=abs(y-pred)) |> 
#   select(-y,-pred) |>
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()
# 
# 
# post_dat  |> group_by(id,condit,Model,Fit_Method,x) |> 
#   mutate(e2=abs(y-pred),Fit_Method=rename_fm(Fit_Method)) |> 
#   summarise(y1=mean(y), pred1=mean(pred),mean_error=abs(y1-pred1)) |>
#   group_by(id,condit,Model,Fit_Method) |> 
#   summarise(mean_error=mean(mean_error)) |> 
#   group_by(condit,Model,Fit_Method) |>
#   summarise(mean_error=mean(mean_error)) |> 
#   arrange(condit,Fit_Method) |>
#   round_tibble(1) |> pander::pandoc.table()

{{< include ../Discussion.qmd >}}

```{r fig.width=11, fig.height=10}
#| label: fig-htw-best-model-old
#| eval: false
#| include: false
#| fig-cap: Difference in model errors for each participant, with models fit to both train and test data. Positive values favor EXAM, while negative values favor ALM. 
#| fig-height: 10
#| fig-width: 11
best_id_x <- post_dat_avg |> group_by(id,condit,Model,Fit_Method,x) |>
  mutate(e2=(y-pred)) |>
  summarise(y=mean(y), pred=mean(pred), mean_error=mean(e2),abs_me=abs(mean_error)) |>
  group_by(id,condit,Fit_Method,x) |> mutate(best=ifelse(mean_error==min(mean_error),1,0)) |>
  group_by(id,condit,Fit_Method,Model) |> mutate(n_best=sum(best))

best_id <- best_id_x |> group_by(id,condit,Fit_Method,Model) |>
  summarise(mean_error=mean(mean_error), n_best=dplyr::first(n_best),abs_me=mean(abs_me))

lowest_error_model <- best_id %>%
  group_by(id, condit,Fit_Method) %>%
  summarise(Best_Model = Model[which.min(mean_error)],
            Best_Model2 = Model[which.min(abs_me)],
            n_best = n_best[which.min(mean_error)],
            Lowest_error = min(mean_error),
            differential = min(mean_error) - max(mean_error)) %>%
  ungroup()

error_difference <- best_id %>%
  select(id, condit, Model,Fit_Method, mean_error,abs_me) %>%
  pivot_wider(names_from = Model, values_from = c(abs_me,mean_error)) %>%
  mutate(Error_difference = (mean_error_ALM - mean_error_EXAM), abs_error_dif = (abs_me_ALM - abs_me_EXAM))

full_comparison |> filter(Fit_Method=="Test_Train") |> 
  ungroup() |>
  mutate(id = reorder(id, abs_error_dif)) %>%
  ggplot(aes(y=id,x=abs_error_dif,fill=Best_Model2))+
  geom_col()+
  ggh4x::facet_grid2(~condit,axes="all",scales="free_y", independent = "y")+
  labs(fill="Best Model",x="Mean Model Error Difference (ALM - EXAM)",y="Participant")

>>>>

```{r  fig.width=11, fig.height=10}
#| eval: false
#| include: false
#| label: fig-htw-indv-pred
#| fig-cap: Model predictions alongside observed data for a subset of individual participants. A) 3 constant and 3 varied participants fit to both the test and training data. B) 3 constant and 3 varied subjects fit to only the trainign data. 
#| fig-height: 13
#| fig-width: 12

@@
